---
title: "lecture4"
author: "me"
date: "6/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

library(tidyverse)
abreu_batted_balls <- 
  read_csv("http://www.stat.cmu.edu/cmsac/sure/2021/materials/data/xy_examples/abreu_2020_batted_balls.csv")
head(abreu_batted_balls)


```

Pitch type
batted ball type
outcome (categorical)

# Visualizing 1D categorical data -----------------

```{r}

abreu_batted_balls %>% 
  ggplot(aes(x = batted_ball_type)) +
  geom_bar() +
  theme_bw()

```

The statistical transformation for the bar plot is the counting, it "counts" how much of
each of the bar types and maps it to the y axis.

This is displaying the marginal distribution of the categorical variable.

If I want to display this as probabilities then I need to compute this seperately.

```{r}

abreu_batted_balls %>% 
  ggplot(aes(x = batted_ball_type)) +
  geom_bar(aes(y = after_stat(count) / sum(after_stat(count)))) + 
  theme_bw()
  

```


To do what we did above manually

```{r}


abreu_batted_balls %>% 
  group_by(batted_ball_type) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  mutate(total = sum(count), 
         prop = count / total) %>% 
  ggplot(aes(x = batted_ball_type)) +
  geom_bar(aes(y = prop), stat="identity") + 
  theme_bw()


```

This is the probability mass function of the probability type.

How do we compute the standard error of the sampled probability of getting a fly ball

```{r}

abreu_batted_balls %>% 
  group_by(batted_ball_type) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  mutate(total = sum(count), 
         prop = count / total,
         se = sqrt(prop *  (1-prop) / total),
         lower = prop - 2*se,
         upper = prop + 2*se) %>% 
  ggplot(aes(x = batted_ball_type)) +
  geom_bar(aes(y = prop), stat="identity") + 
  geom_errorbar(aes(ymin = lower, ymax = upper), color = "red") + 
  theme_bw()

```


Interpretation of our confidence interval

If i repeated this collection of data, and the proportion of flyballs, and i kept doing this over
and over and over again, and then cosntruct an CI for each sample, then 95% of the CI's calculated will contain the true fixed parameter value. 

95% of the intervals will contain the true value. 


## Rearranging to alphebetical order:

```{r}

abreu_batted_balls %>% 
  group_by(batted_ball_type) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  mutate(total = sum(count), 
         prop = count / total,
         se = sqrt(prop *  (1-prop) / total),
         lower = prop - 2*se,
         upper = prop + 2*se,
         batted_ball_type = fct_reorder(batted_ball_type, prop)) %>% 
  ggplot(aes(x = batted_ball_type)) +
  geom_bar(aes(y = prop), stat="identity") + 
  geom_errorbar(aes(ymin = lower, ymax = upper), color = "red") + 
  theme_bw()

```

# Continuous Variables ---------

Center: mean, median, number, and location of modes
Spread : range(max-min), quantiles, variance (std), etc
Shape : skew vs symmetry, outliers

```{r}

summary(abreu_batted_balls$exit_velocity)

sd(abreu_batted_balls$exit_velocity)

abreu_batted_balls %>% 
  ggplot(aes(y = exit_velocity)) +
  geom_boxplot(aes(x = "")) +
  theme_bw()+
  coord_flip()


```

Boxplots are limited in that we don't know how many modes there are in the dataset. 

It can be helpful but not all that useful in itself. 



# histogram ---------

Taking a variable
Binning it
Counting it

It is an attempt of density estimation. Finding some way of smoothing data of a continuous variable.

```{r}

abreu_batted_balls %>% 
  ggplot(aes(x = exit_velocity)) +
  geom_histogram() +
  theme_bw()

```

The box plots give me no sense of peaks in this histogram. 

Within a particular range of values I've received the count for the certain amount of balls.

Histograms are something you have to tune. You are choosing a statistical transformation


How do I normalize it? 


# Another way to display continuous distributions -----------

```{r}

library(ggbeeswarm)

abreu_batted_balls %>% 
  ggplot(aes(y = exit_velocity)) + 
  geom_beeswarm(aes(x = ""),  cex = 3) + 
  theme_bw() + 
  coord_flip()

```


# Violin Plot -------

Another display that works well with other displays is violin plots


```{r}

abreu_batted_balls %>% 
  ggplot(aes(y = exit_velocity, x = "")) +
  geom_violin() +
  geom_boxplot(width = .2) +
  geom_beeswarm(aes(x = ""),  cex = 3) + 
  theme_bw() +
  coord_flip()


```


# ECDF -----


```{r}

abreu_batted_balls %>% 
   ggplot(aes(x = exit_velocity)) + 
  geom_rug(alpha = 0.7) + 
  stat_ecdf() +
  theme_bw() + 
  geom_hline(yintercept = 0.95, linetype = "dashed", color = "red")

```
















