---
title: "Lectuer5"
author: "me"
date: "6/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Purpose: Intro to 2D Visualizations

```{r echo=FALSE}

library(tidyverse)
abreu_batted_balls <- 
  read_csv("http://www.stat.cmu.edu/cmsac/sure/2021/materials/data/xy_examples/abreu_2020_batted_balls.csv")
head(abreu_batted_balls)

```

# Start Examining data

```{r}

abreu_batted_balls %>% 
  ggplot(aes(x = exit_velocity , y = launch_angle)) +
  geom_point() + 
  theme_bw()

```


# Plotting Pitch Types

```{r}

table(abreu_batted_balls$pitch_type)

```


These type of categorical variables we can change the underlying
order and what label each one is.

```{r}

head(abreu_batted_balls)
## recode the labels for this pitch type
## fct_recode allows me to change the levels of a categorical
abreu_batted_balls <- abreu_batted_balls %>% 
  mutate(pitch_type = fct_recode(pitch_type, "Changeup" = "CH", 
                                 "Breaking Ball" = "CU",
                                 "Fastball" = "FC", 
                                 "Fastball" = "FF",
                                 "Fastball" = "FS",
                                 "Breaking Ball" = "KC",
                                 "Fastball" = "SI",
                                 "Breaking Ball" = "SL"))

abreu_batted_balls <- abreu_batted_balls %>% 
  mutate(new_pitch_type = case_when(
    pitch_type == "CH" ~ "Changeup",
    pitch_type %in% c("CU", "SL") ~ "Breaking ball",
    TRUE ~ "Fastball"
  ))


head(abreu_batted_balls)
```

# Plotting 2-D Categorical Variables



```{r}

abreu_batted_balls %>% 
  ggplot(aes(x = batted_ball_type, 
             fill = pitch_type)) +
  geom_bar() + 
  theme_bw()


```

In this you can see the marginal distributions for the different batted ball types. 


```{r}

abreu_batted_balls %>% 
  ggplot(aes(x = batted_ball_type, 
             fill = pitch_type)) +
  geom_bar(position = "dodge") + 
  theme_bw()
```
The focus on this one is comparing the conditional distributions. 

# Manually calculating the frequencies in R



```{r}

abreu_batted_balls %>% 
  group_by(batted_ball_type, pitch_type) %>% 
  count() # the same as summarize(n = n())

```
What is the joint distribution 

```{r}

abreu_batted_balls %>% 
  group_by(batted_ball_type, pitch_type) %>% 
  summarize(joint_prob = n() / nrow(abreu_batted_balls)) %>% 
  pivot_wider(names_from = batted_ball_type,
             values_from = joint_prob,
             values_fill = 0)

```

You can see now there is no combination of change-up and pop-up. 

```{r}

abreu_batted_balls %>% 
  group_by(batted_ball_type, pitch_type ) %>% 
  summarize(count = n(), 
            joint_prob = count / nrow(abreu_batted_balls)) %>% 
  ggplot(aes(x = batted_ball_type, y = pitch_type)) + 
  geom_tile(aes(fill = count), color = "white") + 
  geom_text(aes(label = round(joint_prob, digits = 2)), color = "white") + 
  scale_fill_viridis_b() + 
  theme_bw()

```
# VCD

```{r}

library(vcd)

mosaic(~ pitch_type + batted_ball_type, 
       data = abreu_batted_balls)

```


# Continuous with Categorical Variables

```{r}

abreu_batted_balls %>% 
  ggplot(aes( y = exit_velocity, x = pitch_type)) + 
  geom_violin() + 
  geom_boxplot(width=.2) + 
  theme_bw()

```
```{r}

abreu_batted_balls %>% 
  ggplot(aes( x = exit_velocity, color = pitch_type)) + 
  stat_ecdf()+
  theme_bw() +
  theme(legend.position = "bottom") + 
  labs(x = "Exit velocity",
       y = "Conditional proportion of batted balls")

```


## Histogram

```{r}
abreu_batted_balls %>% 
  ggplot(aes(x = exit_velocity, fill = pitch_type, color = pitch_type)) +
  geom_histogram(position = "identity", alpha = 0.25) + 
  theme_bw() + 
  theme(legend.position = "bottom")

```

```{r}

abreu_batted_balls %>% 
  ggplot(aes(x = exit_velocity, color = pitch_type)) + 
  geom_freqpoly() + 
  theme_bw() + 
  theme(legend.position = "bottom")


```

# Facet (subplots)

```{r}

abreu_batted_balls %>% 
  ggplot(aes(x = exit_velocity)) + 
  geom_histogram() + 
  theme_bw() + 
  facet_wrap(~ pitch_type, ncol = 2 ) #take my continuous variable and display it based on a categorical variable

```

```{r}
abreu_batted_balls %>% 
  ggplot(aes(x = exit_velocity)) + 
  geom_histogram() + 
  theme_bw() + 
  facet_grid(pitch_type ~ . , margins = TRUE) 

abreu_batted_balls %>% 
  ggplot(aes(x = exit_velocity)) + 
  geom_histogram() + 
  theme_bw() + 
  facet_grid(pitch_type ~ batted_ball_type , margins = TRUE) 

```

```{r}

library(ggthemes)

abreu_batted_balls %>% 
  ggplot(aes(x = pitch_type, fill = batted_ball_type)) + 
  geom_bar() + 
  theme_bw() + 
  facet_wrap(~ outcome, ncol = 5) +
  theme(legend.position = "bottom") + 
  ggthemes::scale_fill_colorblind()
  

```





